### 2021-04-11 input/proposal.tex
reworkd abstract and scoped broadcast

Scoped broadcast is a communication primitive that propagates a
message around its broadcaster within an application-dependant scope.
For instance, a process from Paris could scoped broadcast messages to
all processes in Paris; and processes from other cities would never
deliver such messages. Scoped broadcast allows its implementations to
reduce generated traffic compared to uniform reliable
broadcast~\cite{hadzilacos1994modular, raynal2013distributed}, for
message propagation can stop as soon as messages reach the borders of
the scope.  This section introduces model and notations to provide a
general definition of scoped broadcast in distributed
systems. \TODO{to help understand adaptive scoped broadcast :).}
\TODO{Better transition to the rest of this section.}

Distributed systems such as Edge infrastructures comprise
heterogeneous machines interconnected by communication links. We do
not consider byzantine processes.

\begin{definition}[Fog infrastructure]
  A fog infrastructure is a connected \underline{g}raph $G(V, E)$ of
  \underline{v}ertices $V$ and bidirectional \underline{e}dges $E
  \subseteq V \times V$.  A \underline{p}ath $\pi_{ij}$ from Process
  $i$ to Process $j$ is a sequence of vertices $[i, k_1, k_2, \ldots
    k_n, j]$ with $\forall m: 0\leq m \leq n, \langle \pi_{ij}[m],
  \pi_{ij}[m+1] \rangle \in E$.
\end{definition}

Processes can send messages to their neighboring processes. Processes
can reliably broadcast messages by forwarding delivered
messages~\cite{birman1999bimodal, hadzilacos1994modular,
  nedelec2018causal, raynal2013distributed}. However, uniform reliable
broadcast states that \emph{all} correct processes must deliver
broadcast messages. This proves costly in large scale systems
comprising thousands of processes. Instead, we define \emph{scoped
broadcast} where messages reach only interested processes,
significantly reducing generated traffic. \TODO{Epidemic propagation
  and forwarding are well-aligned with scoped broadcast.}

%% Epidemic
%% dissemination~\cite{birman1999bimodal} constitutes an efficient
%% communication pattern to perform such operation. \TODO{More about
%%   \emph{forwarding}.}

\begin{definition}[\label{def:scoped}\underline{S}coped broad\underline{cast} (\NAMEB)]
  When Process $x$ scoped \underline{b}roadcasts $b_x(m)$ a
  \underline{m}essage $m$, every correct Process $y$ within a scope
  \underline{r}eceives $r_y(m)$ and \underline{d}elivers it
  $d_y(m)$. The scope depends on the \underline{s}tate $\sigma$ of
  each process, the \underline{m}etadata $\mu$ piggybacked by each
  message, and a \underline{p}redicate $\phi$ verified from process to
  process: $b_x(m) \wedge r_y(m) \implies \exists \pi_{xy}: \forall z
  \in \pi_{xy}, \phi(\mu_z, \sigma_z)$.
\end{definition}

This definition highlights the transitive relevance of messages and
encompasses more specific definitions of related
work~\cite{hsiao2005scoped, lue2006scoped, wang2015prodiluvian}. For
instance, a process from Paris could scoped broadcast messages to all
processes in Paris. This requires processes to know the city they are
in. In another instance, a process from Paris could scoped broadcast
messages to all processes in Paris plus neighboring cities. This
requires processes to overload forwarded messages the first time they
reach another city. The predicate checks if messages already reached
two distinct cities before delivery. Similarly to uniform reliable
broadcast, scoped broadcast implementations expose different
trade-offs on space, time, and communication.

Next Section introduces an adaptive scoped broadcast where the state
of each process is dynamic and changes depending on all distinct
scopes that exist in the system.

@@

In this section, we propose to dynamically adapt the scopes of \NAMEB
using \NAMEB itself for the sake of efficiency. \TODO{States of
  processes change, predicates change over time.} At any time, a
process decides whether or not it becomes the center (or
\emph{source}) of a scope (or \emph{partition}). Processes must
eventually rally the partition corresponding to their closest source
despite concurrent operations, and dynamic changes in the network.

\noindent In that regard, we introduce \NAME (stands for
\underline{A}daptive \underline{S}coped broad\underline{cast}), a
wait-free reactive protocol for dynamic logical partitioning that
enables adaptive scoped broadcast in dynamic distributed
systems. \NAME's operation guarantees that processes eventually reach
consistent partitioning by efficiently propagating their state
changes. \NAME's overhead actually depends on its operations and
current partitions in the system. When the system becomes quiescent,
processes eventually converge to their respective partition and do not
require further communication afterward.

\noindent This section starts by introducing consistent partitioning
and its implementation when a process can only become a new source to
create a new partition, i.e., it cannot revoke its self-appointed
status of source. Then, it extends to include the latter behavior: it
introduces the properties that guarantee consistent partitioning in
such context, and it details our implementation called \NAME. Finally,
it provides the complexity analysis of \NAME divided in time, space,
and communication.

@@

At any time, a process can decide to become a source, hence creating a
new partition in the system by executing an \texttt{Add}
operation. This partition includes at least its source plus
neighboring processes that estimate they are closer to this source
than any other source. Such distance (or \emph{weight}) is
application-dependant: in the context of maintaining distributed
indexes, weights are latency.

@@

\begin{definition}[\label{def:partitioning}Partitioning]
  Let $S \subseteq V$ be the set of \underline{s}ources, and $P_s$ be
  the \underline{p}artition including at least Process $s$, each
  process belongs to at most one partition $\forall p,q \in V, \forall
  s_1,s_2 \in S: p \in P_{s_1} \wedge q \in P_{s_2} \implies p \neq q
  \vee s_1 = s_2$, and there exists at least one path $\pi_{ps}$ of
  processes that belong to this partition $\forall q \in \pi_{ps}: q
  \in P_s$.
\end{definition}

Definition~\ref{def:scoped} and Definition~\ref{def:partitioning}
share the transitive relevance of process states. However, we further
constrain the partitioning in order to guarantee the existence of
exactly one consistent partitioning that processes eventually converge
to.

\begin{definition}[\underline{C}onsistent \underline{p}artitioning (CP)]
  Let $W_{xy} = W_{yx}$ be the symmetric \underline{w}eight between
  $x$ and $y$, $\Pi_{xz}$ be the shortest \underline{p}ath from $x$ to
  $z$ the weight of which $|\Pi_{xz}|$ is lower than any other path
  weight, the only consistent partitioning $\mathcal{P}$ is a set of
  partitions $P_{s\in S}$ such that each process belongs to a logical
  partition comprising its closest source: $\forall p \in P_{s_1},
  \nexists P_{s_2}$ such that $|\Pi_{s_1p}| > |\Pi_{s_2p}|$.
\end{definition}

Unfortunately, processes do not share a common global knowledge of the
network state. For processes to reach consistent partitioning
together, a process becoming a source must notify every process that
is closer to it than any other source, using \NAMEB for the sake of
efficiency.

\begin{theorem}[\underline{B}est \underline{e}ventual \underline{f}orwarding $(BEF) \iff CP$]
  Assuming reliable communication links where a correct process $q$
  eventually receives the message $m$ \underline{s}ent to it by a
  process $p$ ($s_{pq}(m) \implies r_{q}(m)$); a static system ($G^t =
  G^{t+1}$); assuming local knowledge only, processes eventually reach
  consistent partitioning if and only if each process eventually
  forwards its best known partition.
\end{theorem}

\TODO{If forwarding defined as send to all neighbors, then it also
  works in dynamic systems where processes join and edges are added,
  for when a new neighbor appears, the processes must forward their
  best.}

\begin{proof}
  \begin{asparadesc}
  \item [$BEF \implies CP$:] When a process $s_1$ becomes a source, it
    belongs to its own partition, for there exists no better partition
    than its own: $\forall p \in V: |\Pi_{s_1 s_1}| < |\Pi_{s_1
      p}|$. It broadcasts such message to its neighbors. Since
    communication channels are reliable, neighboring processes
    eventually receive such notification. In particular, whatever the
    order of received messages, every process $p_1$ -- such that there
    exists no better partition $s_2$ than the received one -- delivers
    and forwards it: $\forall s_2 \in S: |\Pi_{p s_1}| < |\Pi_{p
      s_2}|$. The notification emanating from $s_1$ transitively
    reaches all such processes through shortest paths. Since there
    exists only one best value per process that can never be
    retracted, processes eventually reach consistent
    partitioning. \TODO{Rework this with $s_{pq}$ and $r_{pq}$.}
  \item [$CP \implies BEF$:] \TODO{By contradiction, if there is a
    process $p$ that does not forward, then another process $q$ may
    never gets its best because it needs the weight $W_{pq}$.}
  \end{asparadesc}
\end{proof}

\begin{algorithm}
  \input{input/algoadd.tex}
  \caption{\label{algo:add}Adding a partition by Process $p$.}
\end{algorithm}

@@

Algorithm~\ref{algo:add} shows the instructions that implement such
consistent partitioning when
\begin{inparaenum}[(i)]
\item weights are scalar values,
\item processes only add new partitions to the system,
\item and processes never join nor crash nor leave the system.
\end{inparaenum}
Figure~\ref{fig:add} illustrates its behavior on a system comprising 4
processes $a$, $b$, $c$, and $d$. Process~$a$ and Process~$d$ become
the sources of their partition. They \NAMEB a notification
\underline{a}dd message: $\alpha_a^0$ and $\alpha_d^0$. They
initialize their own state with the lowest possible bound $0$ (see
Line~\ref{line:lowestbound}), and send a message to each of their
neighbors by accumulating the corresponding edge weight (see
Line~\ref{line:accumulator}). In Figure~\ref{fig:addC}, Process~$b$
receives $\alpha_{d}^{1}$. Since it improves its own partition
distance, it keeps it and forwards it to its neighbors. In
Figure~\ref{fig:addD}, Process~$b$ discards $\alpha_{a}^{2}$, for it
does not improve its partition distance. Processes $c$ and $d$ will
never hear of Source~$a$. Ultimately, processes discard last
transiting messages. The system reached consistent partitioning.

@@

At any time, a process can become a source, hence adding a new
partition to the system. This partition eventually includes all
processes that are closer from this new source than any other
else. Processes naturally converge towards their respective best
partition by only piggybacking a monotonically increasing distance in
forwarded messages.

Then, at any time, a source can revoke its self-appointed status of
source, hence deleting its partition from the system, by executing a
\texttt{Del} operation. All processes that belong to this partition
must eventually choose another partition to belong to. In
Figure~\ref{fig:del}, two partitions exist initially: $P_a$ and $P_d$
that respectively include $\{a\}$, and $\{b, c, d\}$. In
Figure~\ref{fig:delA}, Process $a$ deletes its partition. It notifies
all neighboring processes -- here only Process $b$ -- that may belong
to its partition using \NAMEB. Upon receipt, Process $b$ discards the
message $\delta_a$, for the latter does not target the former's
partition. Yet, Process $b$ sends its own partition that may be the
best for Process $a$. Eventually, every processes belong to the same
partition $P_d$.

@@

Unfortunately, stale control information about deleted partitions may
impair the propagation of both
\begin{inparaenum}[(i)]
\item notifications about actual sources, and
\item notifications about other deleted partitions.
\end{inparaenum}
Figure~\ref{fig:problem} highlights such consistency issue with
dynamic partitions, even in contexts where processes have FIFO links,
\ie where processes receive the messages in the order of their
sending. Both Process~$a$ and Process~$d$ add then delete their
partition concurrently. Process~$c$ receives, delivers, and forwards
$\alpha_d^3$ followed by $\delta_d$. In the meantime, Process~$b$
receives, delivers, and forwards $\alpha_a^2$. In
Figure~\ref{fig:problemC}, both Process~$c$ and Process~$d$ deliver
the message about Partition $P_a$, for they did not have any known
partition at receipt time. On the contrary, Process~$b$ delivers
$\alpha_d^1$, for it improves its distance to a known
source. Process~$b$ then blocks Process~$a$'s removal notification
$\delta_a$. It never reaches Process~$c$ nor Process~$d$. Also,
Process~$c$ and Process~$d$ do not deliver any $\alpha_d$ or
$\delta_d$ since they already delivered it. The system converges to an
unconsistent state where some processes assume they belong to a
partition while it does not exist anymore. Naive propagation of
$\alpha$ and $\delta$ messages is insufficient to guarantee consistent
partitioning when each process can add or delete its partition at any
time.

One could solve this issue by always propagating the $\delta$ messages
corresponding to the $\alpha$ messages it propagated before. In
Figure~\ref{fig:problem}, it means that as soon as Process~$b$
forwards $\alpha_a^2$, it assumes that its neighbors Process~$c$ and
Process~$d$ may need the notification of removal $\delta_a$ if it
exists. However, such solution also implies that processes generate
traffic not only related to their current partition, but also related
to partitions they belong to in the past. Instead, we propose to use
the delivery order at each process to detect possible inconsistencies
and solve them. Together, processes eventually remove all stale
control information (transiting messages and local states) of the
system leaving room for propagation of messages about up-to-date
partitions.

%% To identify and solve the issue with dynamic consistent partitioning
%% where protocols enable processes to \texttt{Add} and \texttt{Del} at
%% any time, we need to further constrain the delivery order at each
%% process.

@@

\begin{corollary}[\label{theo:dcp}Dynamic consistent partitioning]
  Consistent partitioning where each process can \texttt{Add} and
  \texttt{Del} its partition at any time requires
  \begin{inparaenum}[(i)]
  \item eventual purging of stale notifications, then
  \item eventual delivery of bests partitions.
  \end{inparaenum}
\end{corollary}

@@

We define a last win order to forbid the delivery of stale information
emanating from a same process. In other terms, the \emph{best}
notification message is always the most recent known of each process.

\begin{definition}[Last win order]
  When a process $p$ broadcasts two messages $b_p(m) \rightarrow
  b_p(m')$, no process $q$ can deliver $m$ if it has delivered $m'$:
  $\not\exists q \in V$ with $d_q(m') \rightarrow d_q(m)$.
\end{definition}

@@

Figure~\ref{fig:proof} depicts the issues with staleness and message
orderings. In Figure~\ref{fig:proofA}, the shortest path from any
source to Process $c$ is $[a, b, c]$. However, Process $b$ still holds
a stale $\alpha_d^{0.5}$ without knowing. When it receives
$\alpha_a^1$, it discards it, for it assumes that downstream processes
are more interested in $\alpha_d^{0.5}$. To reach consistent
partitioning, Process $b$ first needs to purge its current partition
to later accept that of its current actual shortest path:
$\alpha_a^1$.

\noindent Figure~\ref{fig:proofB} shows that removing stale control
information is even more complex. The removal must reach all processes
of the previous shortest path going from $d$ to $b''$. Label I' shows
the most obvious issue where Process $e$ changed partition for a
better but stale $\alpha_f$. Since it can remember its previous
deliveries, it could still forward $\delta_d$ for the sake of
consistency. However, this would lead to every process forwarding
every $\delta$ they ever delivered. Such protocol's overhead would
depend on past partitioning instead of current one. Label III shows
the issue when the blocking partition is already known to be stale at
Process $b'$. Process $b''$ eventually receives $\alpha_f$ from
$e$. However, it cannot deliver it, for it would break last win order.
Process $b'$ may be in an inconsistent state. Label IV shows the
corollary issue: Process $b''$ delivered a message from a process that
may be inconsistent without knowing it.

@@

\begin{algorithm}
  \input{input/algoadddelundo.tex}
  \caption{\label{algo:adddelundo}Dynamic partitioning by Process $p$.}
\end{algorithm}

We extend Algorithm~\ref{algo:add} to enable each source to revoke its
self-appointed status of source and become a simple process again.
Algorithm~\ref{algo:adddelundo} provides the instructions that
implement dynamic consistent partitioning:
\begin{inparaenum}[(i)]
\item \label{algo:most} most of the time, it propagates $\alpha$'s and
  $\delta$'s when the partitions allow it;
\item \label{algo:sometimes} sometimes, it detects possible
  inconsistent partitioning and
\item \label{algo:solves}solves it using propagation trees (as opposed
  to propagation graphs);
\item \label{algo:competition} when required, it triggers competitions
  among neighboring partitions.
\end{inparaenum}
While (\ref{algo:most})-(\ref{algo:solves}) tackle the eventual purging of
stale notifications, (\ref{algo:competition}) tackle the eventual
delivery of the best up-to-date partitions.

To constrain the order of delivery at each process, each process
implements last win order by maintaining a vector of versions that
associates to each known source, or has-been source, with its known
local counter. It enables processes to deliver only up-to-date
$\alpha$'s that may improve their best known partition (see
Line~\ref{line:detectA}), but also to detect possible inconsistent
partitioning (see Line~\ref{line:detectB}) as labeled by III in
Figure~\ref{fig:proofB}. From a global perspective, it ensures
monotonic increase of knowledge despite the non-monotonic number of
sources. In turns, it contributes to convergence and termination, for
corresponding $\alpha$ and $\delta$ cannot follow each other in an
infinite loop.

Implementing the propagation of deletes is no different from the
propagation of adds as hinted in Figure~\ref{fig:proof} with Labels I
and I'. Messages and comparisons include the local counter of the
source. When a process receives a better up-to-date partition, it
delivers and forwards it. When a process receives a delete
notification about its partition, it delivers and forwards it. It is
worth noting that such $\delta$'s apply to the whole partition. They
travel in the propagation graph of their partition using all
communication links, allowing processes to change partition as soon as
possible.

>>>>

Scoped broadcast is a communication primitive that propagates a
message around its broadcaster within an application-dependant scope.
For instance, a \process from Paris could scoped broadcast messages to
all \processes in Paris; and \processes from other cities would never
deliver such messages.
%% Dedicated scoped broadcast implementations
%% outperform uniform reliable broadcast
%% implementations~\cite{hadzilacos1994modular, raynal2013distributed}
%% since the former targets a connected \emph{subset} of the whole
%% network while the latter targets \emph{all} \processes of the
%% distributed system.
Scoped broadcast only targets a connected \emph{subset} of \processes
in the whole distributed system.
%
Distributed systems such as Edge infrastructures comprise
heterogeneous machines interconnected by communication
links. \Processes may crash but are not byzantine.

% \TODO{dynamic graph: superscript with $t$?}

\begin{definition}[Fog infrastructure]
  A fog infrastructure is a connected \underline{g}raph $G(V, E)$ of
  \underline{v}ertices $V$ and bidirectional \underline{e}dges $E
  \subseteq V \times V$.  A \underline{p}ath $\pi_{ij}$ from \Process
  $i$ to \Process $j$ is a sequence of vertices $[i, k_1, k_2, \ldots
    k_n, j]$ with $\forall m: 0\leq m \leq n, \langle \pi_{ij}[m],
  \pi_{ij}[m+1] \rangle \in E$.
\end{definition}

\Processes can reliably communicate through asynchronous message
passing to other known \processes called neighbors. Using epidemic
propagation (or gossiping), \processes can reliably broadcast messages
by forwarding delivered messages from neighbors to
neighbors~\cite{birman1999bimodal, hadzilacos1994modular,
  nedelec2018causal, raynal2013distributed}. However, uniform reliable
broadcast states that \emph{all} correct \processes must deliver all
broadcast messages. This proves costly in large scale systems
comprising thousands of \processes. Instead, we define \emph{scoped
broadcast} where messages reach only an application-dependant subset
of connected \processes, thus significantly reducing the generated
traffic of broadcast.

\begin{definition}[\label{def:scoped}\underline{S}coped broad\underline{cast} (\NAMEB)]
  When \Process $x$ scoped \underline{b}roadcasts $b_x(m)$ a
  \underline{m}essage $m$, every correct \process $y$ within a scope
  \underline{r}eceives $r_y(m)$ and \underline{d}elivers it
  $d_y(m)$. The scope depends on the \underline{s}tate $\sigma$ of
  each \process, the \underline{m}etadata $\mu$ piggybacked by each
  message, and a \underline{p}redicate $\phi$ verified from \process to
  \process: {\small{$b_x(m) \wedge r_y(m) \implies \exists \pi_{xy}: \forall z
  \in \pi_{xy}, \phi(\mu_z, \sigma_z)$}}.
\end{definition}

This definition encompasses more specific definitions of related
work~\cite{hsiao2005scoped, lue2006scoped, wang2015prodiluvian}.  This
definition also highlights that epidemic propagation and scoped
broadcast have well-aligned preoccupations: It highlights the
transitive relevance of messages, thus \processes can stop forwarding
messages as soon as the corresponding predicate becomes unverified.
For instance, a \process from Paris could scoped broadcast messages to
all \processes in Paris. This requires \processes to store and
maintain their city location in their local state. \Processes stop
delivering and forwarding their received messages when they come from
a different city.  In another instance, a \process from Paris could
scoped broadcast messages to all \processes in Paris plus neighboring
cities. This requires \processes to overload forwarded messages the
first time they reach another city. The predicate checks if messages
already reached two distinct cities before delivery. Similarly to
uniform reliable broadcast, scoped broadcast implementations expose
different trade-offs on space, time, and communication.

Next Section introduces an adaptive scoped broadcast where the state
of each \process is dynamic and changes depending on all distinct
scopes that exist in the system.

@@

In this section, we propose to dynamically adapt the scopes of \NAMEB
using \NAMEB itself for the sake of efficiency. \TODO{States of
  \processes change, predicates change over time.} At any time, a
\process decides whether or not it becomes the center (or
\emph{source}) of a scope (or \emph{partition}). \Processes must
eventually rally the partition corresponding to their closest source
despite concurrent operations, and dynamic changes in the network.

\noindent In that regard, we introduce \NAME (stands for
\underline{A}daptive \underline{S}coped broad\underline{cast}), a
wait-free reactive protocol for dynamic logical partitioning that
enables adaptive scoped broadcast in dynamic distributed
systems. \NAME's operation guarantees that \processes eventually reach
consistent partitioning by efficiently propagating their state
changes. \NAME's overhead actually depends on its operations and
current partitions in the system. When the system becomes quiescent,
\processes eventually converge to their respective partition and do not
require further communication afterward.

\noindent This section starts by introducing consistent partitioning
and its implementation when a \process can only become a new source to
create a new partition, i.e., it cannot revoke its self-appointed
status of source. Then, it extends to include the latter behavior: it
introduces the properties that guarantee consistent partitioning in
such context, and it details our implementation called \NAME. Finally,
it provides the complexity analysis of \NAME divided in time, space,
and communication.

@@

At any time, a \process can decide to become a source, hence creating a
new partition in the system by executing an \texttt{Add}
operation. This partition includes at least its source plus
neighboring \processes that estimate they are closer to this source
than any other source. Such distance (or \emph{weight}) is
application-dependant: in the context of maintaining distributed
indexes, weights are latency.

@@

\begin{definition}[\label{def:partitioning}Partitioning]
  Let $S \subseteq V$ be the set of \underline{s}ources, and $P_s$ be
  the \underline{p}artition including at least \Process $s$, each
  \process belongs to at most one partition $\forall p,q \in V, \forall
  s_1,s_2 \in S: p \in P_{s_1} \wedge q \in P_{s_2} \implies p \neq q
  \vee s_1 = s_2$, and there exists at least one path $\pi_{ps}$ of
  \processes that belong to this partition $\forall q \in \pi_{ps}: q
  \in P_s$.
\end{definition}

Definition~\ref{def:scoped} and Definition~\ref{def:partitioning}
share the transitive relevance of \process states. However, we further
constrain the partitioning in order to guarantee the existence of
exactly one consistent partitioning that \processes eventually converge
to.

\begin{definition}[\underline{C}onsistent \underline{p}artitioning (CP)]
  Let $W_{xy} = W_{yx}$ be the symmetric \underline{w}eight between
  $x$ and $y$, $\Pi_{xz}$ be the shortest \underline{p}ath from $x$ to
  $z$ the weight of which $|\Pi_{xz}|$ is lower than any other path
  weight, the only consistent partitioning $\mathcal{P}$ is a set of
  partitions $P_{s\in S}$ such that each \process belongs to a logical
  partition comprising its closest source: $\forall p \in P_{s_1},
  \nexists P_{s_2}$ such that $|\Pi_{s_1p}| > |\Pi_{s_2p}|$.
\end{definition}

Unfortunately, \processes do not share a common global knowledge of the
network state. For \processes to reach consistent partitioning
together, a \process becoming a source must notify every \process that
is closer to it than any other source, using \NAMEB for the sake of
efficiency.

\begin{theorem}[\underline{B}est \underline{e}ventual \underline{f}orwarding $(BEF) \iff CP$]
  Assuming reliable communication links where a correct \process $q$
  eventually receives the message $m$ \underline{s}ent to it by a
  \process $p$ ($s_{pq}(m) \implies r_{q}(m)$); a static system ($G^t =
  G^{t+1}$); assuming local knowledge only, \processes eventually reach
  consistent partitioning if and only if each \process eventually
  forwards its best known partition.
\end{theorem}

\TODO{If forwarding defined as send to all neighbors, then it also
  works in dynamic systems where \processes join and edges are added,
  for when a new neighbor appears, the \processes must forward their
  best.}

\begin{proof}
  \begin{asparadesc}
  \item [$BEF \implies CP$:] When a \process $s_1$ becomes a source, it
    belongs to its own partition, for there exists no better partition
    than its own: $\forall p \in V: |\Pi_{s_1 s_1}| < |\Pi_{s_1
      p}|$. It broadcasts such message to its neighbors. Since
    communication channels are reliable, neighboring \processes
    eventually receive such notification. In particular, whatever the
    order of received messages, every \process $p_1$ -- such that there
    exists no better partition $s_2$ than the received one -- delivers
    and forwards it: $\forall s_2 \in S: |\Pi_{p s_1}| < |\Pi_{p
      s_2}|$. The notification emanating from $s_1$ transitively
    reaches all such \processes through shortest paths. Since there
    exists only one best value per \process that can never be
    retracted, \processes eventually reach consistent
    partitioning. \TODO{Rework this with $s_{pq}$ and $r_{pq}$.}
  \item [$CP \implies BEF$:] \TODO{By contradiction, if there is a
    \process $p$ that does not forward, then another \process $q$ may
    never gets its best because it needs the weight $W_{pq}$.}
  \end{asparadesc}
\end{proof}

\begin{algorithm}
  \input{input/algoadd.tex}
  \caption{\label{algo:add}Adding a partition by \Process $p$.}
\end{algorithm}

@@

Algorithm~\ref{algo:add} shows the instructions that implement such
consistent partitioning when
\begin{inparaenum}[(i)]
\item weights are scalar values,
\item \processes only add new partitions to the system,
\item and \processes never join nor crash nor leave the system.
\end{inparaenum}
Figure~\ref{fig:add} illustrates its behavior on a system comprising 4
\processes $a$, $b$, $c$, and $d$. \Process~$a$ and \Process~$d$ become
the sources of their partition. They \NAMEB a notification
\underline{a}dd message: $\alpha_a^0$ and $\alpha_d^0$. They
initialize their own state with the lowest possible bound $0$ (see
Line~\ref{line:lowestbound}), and send a message to each of their
neighbors by accumulating the corresponding edge weight (see
Line~\ref{line:accumulator}). In Figure~\ref{fig:addC}, \Process~$b$
receives $\alpha_{d}^{1}$. Since it improves its own partition
distance, it keeps it and forwards it to its neighbors. In
Figure~\ref{fig:addD}, \Process~$b$ discards $\alpha_{a}^{2}$, for it
does not improve its partition distance. \Processes $c$ and $d$ will
never hear of Source~$a$. Ultimately, \processes discard last
transiting messages. The system reached consistent partitioning.

@@

At any time, a \process can become a source, hence adding a new
partition to the system. This partition eventually includes all
\processes that are closer from this new source than any other
else. \Processes naturally converge towards their respective best
partition by only piggybacking a monotonically increasing distance in
forwarded messages.

Then, at any time, a source can revoke its self-appointed status of
source, hence deleting its partition from the system, by executing a
\texttt{Del} operation. All \processes that belong to this partition
must eventually choose another partition to belong to. In
Figure~\ref{fig:del}, two partitions exist initially: $P_a$ and $P_d$
that respectively include $\{a\}$, and $\{b, c, d\}$. In
Figure~\ref{fig:delA}, \Process $a$ deletes its partition. It notifies
all neighboring \processes -- here only \Process $b$ -- that may belong
to its partition using \NAMEB. Upon receipt, \Process $b$ discards the
message $\delta_a$, for the latter does not target the former's
partition. Yet, \Process $b$ sends its own partition that may be the
best for \Process $a$. Eventually, every \processes belong to the same
partition $P_d$.

@@

Unfortunately, stale control information about deleted partitions may
impair the propagation of both
\begin{inparaenum}[(i)]
\item notifications about actual sources, and
\item notifications about other deleted partitions.
\end{inparaenum}
Figure~\ref{fig:problem} highlights such consistency issue with
dynamic partitions, even in contexts where \processes have FIFO links,
\ie where \processes receive the messages in the order of their
sending. Both \Process~$a$ and \Process~$d$ add then delete their
partition concurrently. \Process~$c$ receives, delivers, and forwards
$\alpha_d^3$ followed by $\delta_d$. In the meantime, \Process~$b$
receives, delivers, and forwards $\alpha_a^2$. In
Figure~\ref{fig:problemC}, both \Process~$c$ and \Process~$d$ deliver
the message about Partition $P_a$, for they did not have any known
partition at receipt time. On the contrary, \Process~$b$ delivers
$\alpha_d^1$, for it improves its distance to a known
source. \Process~$b$ then blocks \Process~$a$'s removal notification
$\delta_a$. It never reaches \Process~$c$ nor \Process~$d$. Also,
\Process~$c$ and \Process~$d$ do not deliver any $\alpha_d$ or
$\delta_d$ since they already delivered it. The system converges to an
unconsistent state where some \processes assume they belong to a
partition while it does not exist anymore. Naive propagation of
$\alpha$ and $\delta$ messages is insufficient to guarantee consistent
partitioning when each \process can add or delete its partition at any
time.

One could solve this issue by always propagating the $\delta$ messages
corresponding to the $\alpha$ messages it propagated before. In
Figure~\ref{fig:problem}, it means that as soon as \Process~$b$
forwards $\alpha_a^2$, it assumes that its neighbors \Process~$c$ and
\Process~$d$ may need the notification of removal $\delta_a$ if it
exists. However, such solution also implies that \processes generate
traffic not only related to their current partition, but also related
to partitions they belong to in the past. Instead, we propose to use
the delivery order at each \process to detect possible inconsistencies
and solve them. Together, \processes eventually remove all stale
control information (transiting messages and local states) of the
system leaving room for propagation of messages about up-to-date
partitions.

%% To identify and solve the issue with dynamic consistent partitioning
%% where protocols enable \processes to \texttt{Add} and \texttt{Del} at
%% any time, we need to further constrain the delivery order at each
%% \process.

@@

\begin{corollary}[\label{theo:dcp}Dynamic consistent partitioning]
  Consistent partitioning where each \process can \texttt{Add} and
  \texttt{Del} its partition at any time requires
  \begin{inparaenum}[(i)]
  \item eventual purging of stale notifications, then
  \item eventual delivery of bests partitions.
  \end{inparaenum}
\end{corollary}

@@

We define a last win order to forbid the delivery of stale information
emanating from a same \process. In other terms, the \emph{best}
notification message is always the most recent known of each \process.

\begin{definition}[Last win order]
  When a \process $p$ broadcasts two messages $b_p(m) \rightarrow
  b_p(m')$, no \process $q$ can deliver $m$ if it has delivered $m'$:
  $\not\exists q \in V$ with $d_q(m') \rightarrow d_q(m)$.
\end{definition}

@@

Figure~\ref{fig:proof} depicts the issues with staleness and message
orderings. In Figure~\ref{fig:proofA}, the shortest path from any
source to \Process $c$ is $[a, b, c]$. However, \Process $b$ still holds
a stale $\alpha_d^{0.5}$ without knowing. When it receives
$\alpha_a^1$, it discards it, for it assumes that downstream \processes
are more interested in $\alpha_d^{0.5}$. To reach consistent
partitioning, \Process $b$ first needs to purge its current partition
to later accept that of its current actual shortest path:
$\alpha_a^1$.

\noindent Figure~\ref{fig:proofB} shows that removing stale control
information is even more complex. The removal must reach all \processes
of the previous shortest path going from $d$ to $b''$. Label I' shows
the most obvious issue where \Process $e$ changed partition for a
better but stale $\alpha_f$. Since it can remember its previous
deliveries, it could still forward $\delta_d$ for the sake of
consistency. However, this would lead to every \process forwarding
every $\delta$ they ever delivered. Such protocol's overhead would
depend on past partitioning instead of current one. Label III shows
the issue when the blocking partition is already known to be stale at
\Process $b'$. \Process $b''$ eventually receives $\alpha_f$ from
$e$. However, it cannot deliver it, for it would break last win order.
\Process $b'$ may be in an inconsistent state. Label IV shows the
corollary issue: \Process $b''$ delivered a message from a \process that
may be inconsistent without knowing it.

@@

\begin{algorithm}
  \input{input/algoadddelundo.tex}
  \caption{\label{algo:adddelundo}Dynamic partitioning by \Process $p$.}
\end{algorithm}

We extend Algorithm~\ref{algo:add} to enable each source to revoke its
self-appointed status of source and become a simple \process again.
Algorithm~\ref{algo:adddelundo} provides the instructions that
implement dynamic consistent partitioning:
\begin{inparaenum}[(i)]
\item \label{algo:most} most of the time, it propagates $\alpha$'s and
  $\delta$'s when the partitions allow it;
\item \label{algo:sometimes} sometimes, it detects possible
  inconsistent partitioning and
\item \label{algo:solves}solves it using propagation trees (as opposed
  to propagation graphs);
\item \label{algo:competition} when required, it triggers competitions
  among neighboring partitions.
\end{inparaenum}
While (\ref{algo:most})-(\ref{algo:solves}) tackle the eventual purging of
stale notifications, (\ref{algo:competition}) tackle the eventual
delivery of the best up-to-date partitions.

To constrain the order of delivery at each \process, each \process
implements last win order by maintaining a vector of versions that
associates to each known source, or has-been source, with its known
local counter. It enables \processes to deliver only up-to-date
$\alpha$'s that may improve their best known partition (see
Line~\ref{line:detectA}), but also to detect possible inconsistent
partitioning (see Line~\ref{line:detectB}) as labeled by III in
Figure~\ref{fig:proofB}. From a global perspective, it ensures
monotonic increase of knowledge despite the non-monotonic number of
sources. In turns, it contributes to convergence and termination, for
corresponding $\alpha$ and $\delta$ cannot follow each other in an
infinite loop.

Implementing the propagation of deletes is no different from the
propagation of adds as hinted in Figure~\ref{fig:proof} with Labels I
and I'. Messages and comparisons include the local counter of the
source. When a \process receives a better up-to-date partition, it
delivers and forwards it. When a \process receives a delete
notification about its partition, it delivers and forwards it. It is
worth noting that such $\delta$'s apply to the whole partition. They
travel in the propagation graph of their partition using all
communication links, allowing \processes to change partition as soon as
possible.

Vectors of versions also allow each \process to detect possible last
win order violations. Upon possible inconsistency, the detecting
\process must remove all $\alpha$ messages originating from it, as
labeled by IV in Figure~\ref{fig:proofB}. Contrarily to delete
operations, this applies to a sub-partition of the current partition,
and cannot rely on local counters to
operate. Algorithm~\ref{algo:adddelundo} implements this behavior by
piggybacking paths in messages. Paths guarantee that such $\delta$
messages apply to all and only $\alpha$'s that followed an identical
path. $\delta$'s propagate in the propagation tree rooted at the
detecting \process. Local states of \processes remain unchanged, but
paths in messages monotonically grow. In turns, $\alpha$ and $\delta$
cannot follow each other in an infinite loop, for a \process knows when
it already received a forwarded a message (see
Lines~\ref{line:notloopingA} and \ref{line:notloopingB}).
